---
title: 'Practical machine learning - Wearables sensors analysis '
author: "Eran Shlomo"
date: "February 15, 2017"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

In this project we will train predictor to identify activity type, based on wearable sensor information collected.
To run the code you will need the datasets,The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


## requirements 
1.Download the datasets into code execution folder. 
2.Install rpart,e1071,randomForest and caret packages. 

## Preparing the data

Setting working dir and loading: 

```{r , warning=FALSE}
library(rpart)
library(caret)
library(e1071)
library(randomForest)
#uncomment if running outside of r markdown 
#this.dir <- dirname(parent.frame(2)$ofile)
#setwd(this.dir)
org_testing = read.csv("./pml-testing.csv",na.strings=c("NA","#DIV/0!",""))
org_training = read.csv("./pml-training.csv",na.strings=c("NA","#DIV/0!",""))
```
Lets split the training data into training/validation (testing)
```{r}
inTrain = createDataPartition(org_training$classe, p = 3/4, list=FALSE)
training = org_training[ inTrain,]
testing = org_training[ -inTrain,]
```
Now lets clean the data:
1. Remove zero variance variables (A LOT OF NA in the dataset)
2. remove id col 
3. clear mostly NA fields (creates a lot of un explained variance)
```{r}
zvars <- nearZeroVar(training)#find  meaningless 
training <- training[,-zvars] #remove meaningless
training <- training[c(-1)] #remove id col

trainingTemp <- training
for(i in 1:length(training)) { #for every column in the training dataset
  if( sum( is.na( training[, i] ) ) /nrow(training) >= .6 ) { 
    for(j in 1:length(trainingTemp)) {
      if( length( grep(names(training[i]), names(trainingTemp)[j]) ) ==1)  { #if the columns are the same:
        trainingTemp <- trainingTemp[ , -j] #Remove that column
      }   
    } 
  }
}
training<-trainingTemp
clean_cols <- colnames(training)
clean_cols_no_classe<-clean_cols[-which(clean_cols %in% c("classe"))]
org_testing<-org_testing[,clean_cols_no_classe]
testing <- testing[,clean_cols]
```

And finally alinging classes and levels:
Classes:
```{r}
for (i in 1:length(org_testing) ) {
  for(j in 1:length(training)) {
    if (names(org_testing[i])==names(training[j]))
    {
      class(org_testing[i]) <- class(training[j])
    }
    #if( length( grep(names(training[i]), names(names(training[2]))[j]) ) ==1)  {
    #  class(org_testing[j]) <- class(training[i])
    #}      
  }      
}
```
levels:
```{r}
common <- intersect(names(training), names(org_testing))
for (p in common)
{ 
  if (class(training[[p]]) == "factor") 
  { 
    levels(org_testing[[p]]) <- levels(training[[p]]) 
  } 
} 
```
# Lets test some prediction models

## Recursive Partitioning and Regression Trees

```{r}
regressionTreeModel <- rpart(classe ~ ., data=training, method="class")
regressionTreePrediction <- predict(regressionTreeModel, testing, type = "class")
confusionMatrix(regressionTreePrediction, testing$classe)
```

## Support vector machine - linear kernel

```{r}
svmModel<-svm(classe~.,data=training,kernel="linear")
svmPrediction <- predict(svmModel, testing)
accuracy=sum(svmPrediction==testing$classe)/dim(testing)[[1]]
print(accuracy)
```


## Support vector machine - radial kernel

```{r}
svmModel<-svm(classe~.,data=training,kernel="radial")
svmPrediction <- predict(svmModel, testing)
accuracy=sum(svmPrediction==testing$classe)/dim(testing)[[1]]
print(accuracy)
```

## Random forest trees

```{r}
randomForestModel <- randomForest(classe ~. , data=training,na.action=na.exclude)
randomForestPrediction <- predict(randomForestModel, testing, type = "class")
confusionMatrix(randomForestPrediction, testing$classe)
```

# Results

## Selected model 
With 99% accuracy, random forest was selected as preffered regression. 

## Running test set,predicting lables 

```{r}
randomForestTestPredict <- predict(randomForestModel, org_testing, type = "class")
print(randomForestTestPredict)
```


